From Reversible Programs to \\ Univalent Universes and Back

### 2017-06-30 upi.tex
add a sentence about the canonical form being constructive.  Add an edcomm about the inconsistent notation.

Having illustrated the general flavor of the $\Pi$ family of
languages, we present in full detail a small $\Pi$-based language
which we will use in the formalization in the rest of the paper. The
language below is the restriction of $\Pi$ to the case of just one
type, the type of booleans:

\[\def\arraystretch{0.8}\begin{array}{l@{\quad}rclrl}
(\textit{Types}) & \tau &::=& \bt \\
\\
(\textit{Terms}) &  v &::=& \fc &:& \bt \\
              && \alt & \tc &:& \bt \\
\\
 (\textit{1-combinators}) &  c &::=& \id &:& \tau \iso \tau \\
               && \alt & \swap &:& \bt \iso \bt \\
               && \alt & ! &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_1) \\
               && \alt & \circ &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_3) \to (\tau_1 \iso \tau_3)  \\
\\
(\textit{2-combinators}) & \alpha &::=& \id &:& c \isotwo c \\
            && \alt & \idlc &:& \compc{\id}{c} \isotwo c \\
            && \alt & \idrc &:& \compc{c}{\id} \isotwo c \\
            && \alt & \invl &:& \compc{c\;}{\;\invc{c}} \isotwo \id \\
            && \alt & \invr &:& \compc{\invc{c}}{c} \isotwo \id \\
            && \alt & \rho &:& \swap \circ \swap \isotwo \id \\
            && \alt & \assocc &:& 
                                  \compc{(\compc{c_1}{c_2})}{c_3} \isotwo \compc{c_1}{(\compc{c_2}{c_3})} \\
            && \alt & \odot &:& (c_1 \isotwo c_1') \to (c_2 \isotwo c_2') \to 
                             (\compc{c_1}{c_2} \isotwo \compc{c_1'}{c_2'}) \\
            && \alt & !! &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_1) \\
            && \alt & \bullet &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_3) \to (c_1 \isotwo c_3)
\end{array}\]

@@

It is relatively easy to think of a few sound program transformations
to include in the category $\alpha$ of 2-combinators. The set above
however is \emph{complete}.

>>>>

Having illustrated the general flavor of the $\Pi$ family of
languages, we present in full detail a small $\Pi$-based language
which we will use in the formalization in the rest of the paper. The
language is the restriction of $\Pi$ to the case of just one
type, the type of booleans:

\jacques{the code above uses $\odot$ for 1-composition, 
$\boxdot$ for parallel 2-composition of $\odot$, while the
code below uses $\circ$ and $\odot$ respectively, which is 
quite confusing.  We should pick one notation.}
\[\def\arraystretch{0.8}\begin{array}{l@{\quad}rclrl}
(\textit{Types}) & \tau &::=& \bt \\
\\
(\textit{Terms}) &  v &::=& \fc &:& \bt \\
              && \alt & \tc &:& \bt \\
\\
 (\textit{1-combinators}) &  c &::=& \id &:& \tau \iso \tau \\
               && \alt & \swap &:& \bt \iso \bt \\
               && \alt & ! &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_1) \\
               && \alt & \circ &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_3) \to (\tau_1 \iso \tau_3)  \\
\\
(\textit{2-combinators}) & \alpha &::=& \id &:& c \isotwo c \\
            && \alt & \idlc &:& \compc{\id}{c} \isotwo c \\
            && \alt & \idrc &:& \compc{c}{\id} \isotwo c \\
            && \alt & \invl &:& \compc{c\;}{\;\invc{c}} \isotwo \id \\
            && \alt & \invr &:& \compc{\invc{c}}{c} \isotwo \id \\
            && \alt & \rho &:& \swap \circ \swap \isotwo \id \\
            && \alt & \assocc &:& 
                                  \compc{(\compc{c_1}{c_2})}{c_3} \isotwo \compc{c_1}{(\compc{c_2}{c_3})} \\
            && \alt & \odot &:& (c_1 \isotwo c_1') \to (c_2 \isotwo c_2') \to 
                             (\compc{c_1}{c_2} \isotwo \compc{c_1'}{c_2'}) \\
            && \alt & !! &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_1) \\
            && \alt & \bullet &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_3) \to (c_1 \isotwo c_3)
\end{array}\]

@@

It is relatively easy to think of a few sound program transformations
to include in the category $\alpha$ of 2-combinators. The set above
however is \emph{complete}. We also have canonical forms (constructively).
